Computation and comparison of the stable Northeastern US marine boundary layer

### 2020-12-04 Paper/main.tex
Added Myra edits

\begin{abstract}
  In this work we investigate the behavior of stable marine boundary
  layers located near the coast of the Northeastern United States.
  Using the ExaWind LES codes, three stable atmospheric conditions
  were chosen to match the Cape Wind measurements of Archer \emph{et
    al.} with wind speeds of 5 m/s, 10 m/s, and 15 m/s at the 20m
  measurement height.  The behavior of the stable boundary layers,
  including mean flow quantities and turbulent statistics, are
  examined and compared to previous computations of the neutral and
  unstable offshore boundary layer at the same location.  This study
  also examines the domain and mesh requirements necessary to capture
  the turbulent scales for the stable offshore boundary layers.
  Finally, we compare solutions computed using both Nalu-Wind and
  AMR-Wind solvers, and report on their relative accuracy and
  performance in this study.

@@

\section{Nomenclature}
\fxnote{Update symbols}
{\renewcommand\arraystretch{1.0}
  \noindent\begin{longtable*}{@{}l @{\quad=\quad} l@{}}
ABL       & Atmospheric Boundary Layer \\
$\alpha$  & Wind shear exponent  \\
$f$       & Frequency \\
$f_{max}$  & Maximum resolvable frequency due to mesh limits \\
$f_{Ny}$   & Nyquist frequency \\
$L$       & Turbulent integral length scale \\
$L_{Ob}$   & Obukhov length scale \\
LES       & Large Eddy Simulation \\
$u_\tau$   & Friction velocity    \\
$\rho$    & Density \\
$\rho_\circ$ & Reference density  \\
$R_{ij}$   & Two point correlation tensor \\
SBL       & Stable Boundary Layer \\
$\theta$  & Potential temperature \\
$TI$      & Turbulence intensity \\
$TKE$     & Turbulent kinetic energy
\end{longtable*}}

@@

In 2000, Kosovic and Curry \cite{Kosovic2000} used an LES simulation
with a nonlinear SGS model containing backscatter to favorably compare
to the Beaufort Sea Arctic Stratus Experiment (BASE) measurements,
showing that the SBL could reach a quasi steady-state. They also
showed good agreement with the Nieustadt analytical model. In the same
year, Saiki et al. \cite{Saiki2000} simulated an SBL with
a strong capping inversion at z=500m, producing reasonable surface layer profiles and
identifying a need for SBL specific SGS models. Beare and MacVean \cite{Beare2004}
presented the first grid-focused work on LES of the SBL, using grids with deltas
between 2m and 12.5m, finding significant variation between grid sizes. The inclusion
of a backscatter model resulted in significant variation at a 6.25m resolution, but
very little at a 2m resolution. Basu et al. \cite{Basu2005} published a set of
turbulence statistics, measurements, and LES simulations in support
of the viability of Nieustadt's local scaling hypothesis under very stable
conditions. Jimenez and Cuxart \cite{Jimenez2005} set out to quantify the range
for which standard SGS models, using the Kolmogorov theory for dissipation, were
valid for increasingly stable flows. They found the model to perform adequately
for weakly stable, and some moderately stable flow regimes.

@@

More recent work in LES of the SBL includes Matheou \cite{Matheou2016} who
explored a large parameter space using the Smagorinsky-Lilly SGS
closure, as well as Sullivan et al.\ \cite{sullivan2016turbulent}. Several researchers
have applied SBL simulation in the wind energy space \cite{Aitken2014, Abkar2015, Ghaisas2017}.
In 2020, Wurps et al. \cite{Wurps2020} provided a rigorous grid resolution study for the SBL,
looking at mean flow, turbulence variables, and two-point correlations, using a dynamic
SGS model that allows backscatter. They found grid requirements to be 2.5m, 10m, and 20m
for stable, neutral, and convective boundary layers respectively. They advocate looking
at two point correlations in order to quantify flow structure size, which in turn
provides a basis for evaluating grid resolution. In all of the reviewed literature, SBL LES
has been run in the onshore context only, a complete comparison
including offshore stable ABL conditions has yet to be completed.

\subsection{Objectives}
In the current study on stable marine boundary layers, we aim to build
on the previous work in the literature and complete several
objectives.  First, a series of stable boundary layer conditions will
be computed using the Exawind solvers which match the measurements of
Archer et al. \cite{archer2016predominance}.  The behavior of the
stable boundary layers, including mean flow quantities and turbulent
statistics, are examined and compared to the neutral and unstable
cases computed in \cite{cheung2020large}.  Secondly, we examine the
domain and mesh requirements necessary to capture the turbulent scales
for the stable offshore boundary layers.  Finally, two LES solvers
within the Exawind project are compared in this study: the
unstructured, incompressible Navier-Stokes solver Nalu-Wind and the
structured, adaptive mesh refinement solver AMR-Wind.  Both codes are
used to calculate the same SBL, and the relative differences between
the codes are reported.

@@

In addition to these simulations, we also examined the domain and mesh
requirements necessary for the stable offshore boundary layers
considered here.  Based on the calculated turbulent integral
lengthscale, horizontal domain sizes of 750m $\times$ 750m were found
to be necessary for these simulations.  Various mesh resolutions were
also considered for the stable runs, and it was determined that the
finest resolution of $dx$=2.5m was able to resolve the longitudinal
length scales and highest temporal frequencies, but inadequate for
resolving the turbulence in the lateral direction.  Further study is
warranted to comprehensively determine the impact of stable
stratification on the required mesh resolution.

>>>>

\begin{abstract}
  In this work we investigate the behavior of stable marine boundary
  layers located near the coast of the Northeastern United States.
  Using the ExaWind LES codes, three stable atmospheric conditions
  were chosen to match the Cape Wind measurements of Archer \emph{et
    al.} with wind speeds of 5 m/s, 10 m/s, and 15 m/s at the 20 m
  measurement height.  The behavior of the stable boundary layers,
  including mean flow quantities and turbulent statistics, are
  examined and compared to previous computations of the neutral and
  unstable offshore boundary layer at the same location.  This study
  also examines the domain and mesh requirements necessary to capture
  the turbulent scales for the stable offshore boundary layers.
  Finally, we compare solutions computed using both Nalu-Wind and
  AMR-Wind solvers, and report on their relative accuracy and
  performance in this study.

@@

\section{Nomenclature}
\fxnote{Update symbols}
{\renewcommand\arraystretch{1.0}
  \noindent\begin{longtable*}{@{}l @{\quad=\quad} l@{}}
ABL       & Atmospheric Boundary Layer \\
$\alpha$  & Wind shear exponent  \\
$f$       & Frequency \\
$f_{max}$  & Maximum resolvable frequency due to mesh limits \\
$f_{Ny}$   & Nyquist frequency \\
$L$       & Turbulent integral length scale \\
$L_{Ob}$   & Obukhov length scale \\
LES       & Large Eddy Simulation \\
$u_\tau$   & Friction velocity    \\
$\rho$    & Density \\
$\rho_\circ$ & Reference density  \\
$R_{ij}$   & Two point correlation tensor \\
SBL       & Stable Boundary Layer \\
$\theta$  & Potential temperature \\
$TI$      & Turbulence intensity \\
$TKE$     & Turbulent kinetic energy \\
$\overline{U}_{horiz}$ & Time averaged horizontal velocity
\end{longtable*}}

@@

In 2000, Kosovic and Curry \cite{Kosovic2000} used an LES simulation
with a nonlinear SGS model containing backscatter to favorably compare
to the Beaufort Sea Arctic Stratus Experiment (BASE) measurements,
showing that the SBL could reach a quasi steady-state. They also
showed good agreement with the Nieustadt analytical model. In the same
year, Saiki et al. \cite{Saiki2000} simulated an SBL with
a strong capping inversion at z=500 m, producing reasonable surface layer profiles and
identifying a need for SBL specific SGS models. Beare and MacVean \cite{Beare2004}
presented the first grid-focused work on LES of the SBL, using grids with deltas
between 2m and 12.5m, finding significant variation between grid sizes. The inclusion
of a backscatter model resulted in significant variation at a 6.25m resolution, but
very little at a 2m resolution. Basu et al. \cite{Basu2005} published a set of
turbulence statistics, measurements, and LES simulations in support
of the viability of Nieustadt's local scaling hypothesis under very stable
conditions. Jimenez and Cuxart \cite{Jimenez2005} set out to quantify the range
for which standard SGS models, using the Kolmogorov theory for dissipation, were
valid for increasingly stable flows. They found the model to perform adequately
for weakly stable, and some moderately stable flow regimes.

@@

More recent work in LES of the SBL includes Matheou \cite{Matheou2016} who
explored a large parameter space using the Smagorinsky-Lilly SGS
closure, as well as Sullivan et al.\ \cite{sullivan2016turbulent}. Several researchers
have applied SBL simulation in the wind energy space \cite{Aitken2014, Abkar2015, Ghaisas2017}.
In 2020, Wurps et al. \cite{Wurps2020} provided a rigorous grid resolution study for the SBL,
looking at mean flow, turbulence variables, and two-point correlations, using a dynamic
SGS model that allows backscatter. They found grid requirements to be 2.5 m, 10 m, and 20 m
for stable, neutral, and convective boundary layers respectively. They advocate looking
at two point correlations in order to quantify flow structure size, which in turn
provides a basis for evaluating grid resolution. In all of the reviewed literature, SBL LES
has been run in the onshore context only, a complete comparison
including offshore stable ABL conditions has yet to be completed.

\subsection{Objectives}
\fxnote{fix order of objectives}
In the current study on stable marine boundary layers, we aim to build
on the previous work in the literature and complete several
objectives.  First, a series of stable boundary layer conditions will
be computed using the ExaWind solvers which match the measurements of
Archer et al. \cite{archer2016predominance}.  The behavior of the
stable boundary layers, including mean flow quantities and turbulent
statistics, are examined and compared to the neutral and unstable
cases computed in \cite{cheung2020large}.  Secondly, we examine the
domain and mesh requirements necessary to capture the turbulent scales
for the stable offshore boundary layers.  Finally, two LES solvers
within the ExaWind project are compared in this study: the
unstructured, incompressible Navier-Stokes solver Nalu-Wind and the
structured, adaptive mesh refinement solver AMR-Wind.  Both codes are
used to calculate the same SBL, and the relative differences between
the codes are reported.

@@

In addition to these simulations, we also examined the domain and mesh
requirements necessary for the stable offshore boundary layers
considered here.  Based on the calculated turbulent integral
lengthscale, horizontal domain sizes of 750 m $\times$ 750 m were found
to be necessary for these simulations.  Various mesh resolutions were
also considered for the stable runs, and it was determined that the
finest resolution of $dx$=2.5 m was able to resolve the longitudinal
length scales and highest temporal frequencies, but inadequate for
resolving the turbulence in the lateral direction.  Further study is
warranted to comprehensively determine the impact of stable
stratification on the required mesh resolution.

### 2020-12-04 Paper/methodology.tex
Added Myra edits

Following the work of Cheung et al \cite{cheung2020large}, we use
measurements of the offshore coastal marine boundary layer, collected
by the Cape Wind meteorological tower in Nantucket Sound, as the basis
for this computational study.  The Cape Wind platform collected wind
measurements at 20m, 41m, and 60m above the mean water level, along
with temperature and barometric measurements from the years 2003-2011.
From the observations, Archer et al.\cite{archer2016predominance}
found that the marine boundary layer is predominantly unstable, with
61\% of conditions classified as unstable, versus 21\% neutral and
18\% stable.  The stratification of the marine boundary layer, as
determined by the Obukhov length, also had a large impact on the wind
speed profile, with flatter, non-logarithmic profiles seen during
unstable conditions.

From the measured distribution of atmospheric stabilities, turbulent
kinetic energies (TKE), and wind speeds from the Cape Wind platform at
z=20m, a stable set of conditions were chosen as targets for this
computational study.  A summary of the conditions for all stability
classes is given in table \ref{tab:CapeWindMeasurements}. In this work we
focus on the three stable atmospheric conditions at wind speeds of
5m/s, 10m/s, and 15m/s, while the neutral and unstable conditions
where studied in previous work \cite{cheung2020large}.

To maintain consistency with the measured data, the turbulence
intensity (TI) is calculated using the TKE as
\begin{equation}
  \textrm{TI} =
  \frac{\sqrt{\frac{2}{3}\times\textrm{TKE}}}{\overline{U}_{horiz}} =
  \frac{\sqrt{\frac{1}{3}\left( \overline{u'u'} + \overline{v'v'} + \overline{w'w'}
      \right)}}{\overline{U}_{horiz}}
\end{equation}
The averaged wind speeds were enforced at the measurement height of
20m, and the applied wind directions were consistent with the
predominant stable direction of 225 degrees southwest.

\begin{table}
\caption{\label{tab:CapeWindMeasurements} Measured conditions at Cape
  Wind.  The stable atmospheric conditions used in this study are
  highlighted in bold below.} \centering
\begin{tabular}{cccc}
  \hline
  Stability    & Wind speed [m/s] & Wind dir [deg] & Turbulence intensity \\
  \hline
  Neutral      & 5                & 225            & 0.055           \\
  Neutral      & 10               & 225            & 0.055           \\
  Neutral      & 15               & 225            & 0.065           \\
  Unstable     & 5                & 315            & 0.080           \\
  Unstable     & 10               & 315            & 0.075           \\
  Unstable     & 15               & 315            & 0.090           \\
  \bf{Stable}  & \bf{5}           & \bf{225}       & \bf{0.045}      \\
  \bf{Stable}  & \bf{10}          & \bf{225}       & \bf{0.050}      \\
  \bf{Stable}  & \bf{15}          & \bf{225}       & \bf{0.060}      \\
\hline
\end{tabular}
\end{table}

@@

Large Eddy Simulations of the stable atmopsheric conditions described in
Table~\ref{tab:CapeWindMeasurements} are performed using the open-source,
ExaWind simulation environment\footnote{\url{https://github.com/exawind}}.
ExaWind is a suite of high-fidelity modeling tools for analyzing complex
microscale atmospheric boundary layer flows and wind farm flows. The present
study uses two different incompressible computational fluid dynamic solvers
available within the ExaWind suite,
Nalu-Wind~\footnote{\url{https://github.com/exawind/nalu-wind}} and
AMR-Wind~\footnote{\url{https://github.com/exawind/amr-wind}}. This section
briefly describes the incompressible Navier-Stokes equations solved,
the specifics of the turbulence model and the wall shear stress
formulation used to model the terrain, and the details of the numerical
discretization in the two codes.

@@

% \fxnote{Do we want a github link reference for amr-wind? exawind?}\\

As part of a suite of physics codes within the open-source project called ExaWind,
AMR-Wind enables more efficient and scalable simulations of wind power plants.
This is because it is built on top of the AMReX software framework which contains
all of the functionality needed to develop massively parallel, block-structured
adaptive mesh refinement (AMR) applications \cite{AMReX_JOSS}.
Since AMR-Wind is limited to Cartesian block-structured grids, it is specialized
to handle ABL physics and wind turbine wakes.
The block-structured grids enable more efficient algorithms such as
Multi-Level Multi-Grid (MLMG) \cite{AMReX_JOSS}
and are well suited for next generation supercomputers that use GPU's.

@@

The computational methodology for the AMR-Wind and Nalu-Wind LES codes
follows practices similar to the previous offshore ABL study \cite{cheung2020large}, with
some modifications to handle the stable stratification.  As in
\cite{cheung2020large}, the domain was a square prism geometry, with
the $x$ and $y$ coordinates aligned in the East and North directions,
respectively.  The domain size and mesh requirements were determined
through a numerical grid study.  As discussed in section
\ref{sec:gridstudy}, the horizontal dimensions of 750m $\times$ 750m
were found to be sufficient to capture any large scale structures in
the boundary layer, and in all cases the vertical dimension of 1000m
was also used.  The grid study in section \ref{sec:gridstudy} also
showed that the stable stratification required greater mesh refinement
to capture the smaller turbulent scales, and a uniform cell size of
2.5m was adopted for the offshore LES computations across both codes.

@@

The initial temperature profile in all cases was a constant 300K until
the specified inversion height of 650m.  The inversion layer thickness
was 100m, and above this, the temperature linearly increased until it
reached 308.75K at the upper boundary.  The initial mean velocity
profile was uniform throughout the domain with superimposed sinusoidal
velocity perturbations of magnitude 1 m/s to promote the development
of turbulence.

\subsubsection{Determination of surface roughness and prescribed temperature }
To match the measured TI conditions given in table
\ref{tab:CapeWindMeasurements}, the surface roughness and prescribed
surface temperature change were adjusted through an initial
trial-and-error process.  The final surface roughness of $z_0$=0.0005m
used for the stable offshore conditions (see table \ref{tab:z0tempparam})
matched the roughness used in the neutral 15m/s and unstable 5m/s and
10m/s cases from the previous study \cite{cheung2020large}.  This
value of surface roughness is consistent with the measurements from
the North Sea \cite{taylor2001dependence}, which found values of $z_0$
ranging from $5 \times 10^{-5}$m to $5\times 10^{-3}$m.

@@

Once the appropriate surface roughness and prescribed temperature were
determined, and the correct mesh requirements known from the study in
section \ref{sec:gridstudy}, the LES computations for the stable 5m/s,
10m/s, and 15m/s ABL cases could be set up and run.  Each of the cases
was run for 15,000 seconds before collecting statistics for another
5,000 seconds.  To maintain a CFL number less than unity during these
runs, a smaller timestep was used for the higher wind speeds (see
table \ref{tab:z0tempparam}).

>>>>

Following the work of Cheung et al \cite{cheung2020large}, we use
measurements of the offshore coastal marine boundary layer, collected
by the Cape Wind meteorological tower in Nantucket Sound, as the basis
for this computational study.  The Cape Wind platform collected wind
measurements at 20 m, 41 m, and 60 m above the mean water level, along
with temperature and barometric measurements from the years 2003-2011.
From the observations, Archer et al.\cite{archer2016predominance}
found that the marine boundary layer is predominantly unstable, with
61\% of conditions were classified as unstable, versus 21\% neutral and
18\% stable.  The stratification of the marine boundary layer, as
determined by the Obukhov length, also had a large impact on the wind
speed profile, with flatter, non-logarithmic profiles seen during
unstable conditions.

From the measured distribution of atmospheric stabilities, turbulent
kinetic energies (TKE), and wind speeds from the Cape Wind platform at
z=20 m, a stable set of conditions were chosen as targets for this
computational study.  A summary of the conditions for all stability
classes is given in table \ref{tab:CapeWindMeasurements}. In this work we
focus on the three stable atmospheric conditions at wind speeds of
5 m/s, 10 m/s, and 15 m/s, while the neutral and unstable conditions
where studied in previous work \cite{cheung2020large}.

To maintain consistency with the measured data, the turbulence
intensity (TI) is calculated using the TKE as
\begin{equation}
  \textrm{TI} =
  \frac{\sqrt{\frac{2}{3}\times\textrm{TKE}}}{\overline{U}_{horiz}} =
  \frac{\sqrt{\frac{1}{3}\left( \overline{u'u'} + \overline{v'v'} + \overline{w'w'}
      \right)}}{\overline{U}_{horiz}}
\end{equation}
The averaged wind speeds were enforced at the measurement height of
20 m, and the applied wind directions were consistent with the
predominant stable direction of 225 degrees southwest.

\begin{table}
\caption{\label{tab:CapeWindMeasurements} Measured conditions at Cape
  Wind.  The stable atmospheric conditions used in this study are
  highlighted in bold below, while the neutral and unstable conditions
  from \cite{cheung2020large} are included for comparison.} \centering
\begin{tabular}{cccc}
  \hline
  Stability    & Wind speed [m/s] & Wind dir [deg] & Turbulence intensity \\
  \hline
  Neutral      & 5                & 225            & 0.055           \\
  Neutral      & 10               & 225            & 0.055           \\
  Neutral      & 15               & 225            & 0.065           \\
  Unstable     & 5                & 315            & 0.080           \\
  Unstable     & 10               & 315            & 0.075           \\
  Unstable     & 15               & 315            & 0.090           \\
  \bf{Stable}  & \bf{5}           & \bf{225}       & \bf{0.045}      \\
  \bf{Stable}  & \bf{10}          & \bf{225}       & \bf{0.050}      \\
  \bf{Stable}  & \bf{15}          & \bf{225}       & \bf{0.060}      \\
\hline
\end{tabular}
\end{table}

@@

Large Eddy Simulations of the stable atmopsheric conditions described in
table~\ref{tab:CapeWindMeasurements} are performed using the open-source,
ExaWind simulation environment\footnote{\url{https://github.com/exawind}}.
ExaWind is a suite of high-fidelity modeling tools for analyzing complex
microscale atmospheric boundary layer flows and wind farm flows. The present
study uses two different incompressible computational fluid dynamic solvers
available within the ExaWind suite,
Nalu-Wind~\footnote{\url{https://github.com/exawind/nalu-wind}} and
AMR-Wind~\footnote{\url{https://github.com/exawind/amr-wind}}. This section
briefly describes the incompressible Navier-Stokes equations solved,
the specifics of the turbulence model and the wall shear stress
formulation used to model the terrain, and the details of the numerical
discretization in the two codes.

@@

\fxnote{Clarify we don't actually use adaptive refinement -- that is future work.}

As part of a suite of physics codes within the open-source project called ExaWind,
AMR-Wind enables more efficient and scalable simulations of wind power plants.
This is because it is built on top of the AMReX software framework which contains
all of the functionality needed to develop massively parallel, block-structured
adaptive mesh refinement (AMR) applications \cite{AMReX_JOSS}.
Since AMR-Wind is limited to Cartesian block-structured grids, it is specialized
to handle ABL physics and wind turbine wakes.
The block-structured grids enable more efficient algorithms such as
Multi-Level Multi-Grid (MLMG) \cite{AMReX_JOSS}
and are well suited for next generation supercomputers that use GPU's.
In the current work, we focus on applying AMR-Wind to ABL computations
using statically refined, uniform meshes, and will examine adaptive
mesh capabilities in future studies.

@@

The computational methodology for the AMR-Wind and Nalu-Wind LES codes
follows practices similar to the previous offshore ABL study \cite{cheung2020large}, with
some modifications to handle the stable stratification.  As in
\cite{cheung2020large}, the domain was a square prism geometry, with
the $x$ and $y$ coordinates aligned in the East and North directions,
respectively.  The domain size and mesh requirements were determined
through a numerical grid study.  As discussed in section
\ref{sec:gridstudy}, the horizontal dimensions of 750 m $\times$ 750 m
were found to be sufficient to capture any large scale structures in
the boundary layer, and in all cases the vertical dimension of 1000 m
was also used.  The grid study in section \ref{sec:gridstudy} also
showed that the stable stratification required greater mesh refinement
to capture the smaller turbulent scales, and a uniform cell size of
2.5m was adopted for the offshore LES computations across both codes.

@@

The initial temperature profile in all cases was a constant 300K until
the specified inversion height of 650 m.  The inversion layer thickness
was 100 m, and above this, the temperature linearly increased until it
reached 308.75K at the upper boundary.  The initial mean velocity
profile was uniform throughout the domain with superimposed sinusoidal
velocity perturbations of magnitude 1 m/s to promote the development
of turbulence.

\subsubsection{Determination of surface roughness and prescribed temperature }
To match the measured TI conditions given in table
\ref{tab:CapeWindMeasurements}, the surface roughness and prescribed
surface temperature change were adjusted through an initial
trial-and-error process.  The final surface roughness of $z_0$=0.0005m
used for the stable offshore conditions (see table \ref{tab:z0tempparam})
matched the roughness used in the neutral 15m/s and unstable 5m/s and
10 m/s cases from the previous study \cite{cheung2020large}.  This
value of surface roughness is consistent with the measurements from
the North Sea \cite{taylor2001dependence}, which found values of $z_0$
ranging from $5 \times 10^{-5}$m to $5\times 10^{-3}$m.

@@

Once the appropriate surface roughness and prescribed temperature were
determined, and the correct mesh requirements known from the study in
section \ref{sec:gridstudy}, the LES computations for the stable 5 m/s,
10 m/s, and 15 m/s ABL cases could be set up and run.  Each of the cases
was run for 15,000 seconds before collecting statistics for another
5,000 seconds.  To maintain a CFL number less than unity during these
runs, a smaller timestep was used for the higher wind speeds (see
table \ref{tab:z0tempparam}).
